---
title: "Introduction to DMAR"
author: "Ken Kelley"
date: "`r format(Sys.Date(), '%B %Y')`"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Introduction to DMAR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width  = 7,
  fig.height = 4.5
)
library(DMAR)
```

## What Is DMAR?

DMAR is a more modern, more general, and greatly expanded reimagining
of the **MBESS** package (Kelley, 2007a, *Journal of Statistical Software*; 2007b, *Behavior Research Methods*). MBESS
itself was originally framed as *Methods for the Behavioral,
Educational, and Social Sciences*, but the methods it implements
became useful far beyond that scope; DMAR ships under a new name to
make that scope expansion explicit. Its emphasis is the methodology
that the substantive disciplines depend on most: confidence intervals
for effect sizes built from noncentral sampling distributions, sample
size planning for both power and accuracy in parameter estimation
(AIPE), reliability, factor analysis, and the critical values used in
classical inference. DMAR is intended to serve as the computational
companion to Maxwell, Delaney, and Kelley's *Designing Experiments
and Analyzing Data* (4th ed., 2027; henceforth **MDK**), and it is
suitable for both instruction and substantive research.

Although the package's origins are in psychology, the questions it
answers are not field-specific. Researchers in biostatistics
(clinical-trial effect size CIs and AIPE-based trial sizing),
organizational behavior and management science (intervention,
training, and program evaluation), information systems (user-experience
A/B comparisons; technology-adoption studies), education (instructional
effectiveness, longitudinal achievement gains), sociology (program
evaluation, attitudinal surveys), and the methodological literature
itself will find tools here for the same three tasks: estimating effect
sizes, attaching confidence intervals that hold their stated coverage,
and planning future
studies that will not be underpowered or underestimated.

What DMAR keeps from MBESS is the philosophy, that is, the same insistence on
*effect sizes with uncertainty quantified*, and on planning studies for
*accurate estimation* rather than only for sufficient power. What DMAR
changes is the implementation: a uniform `snake_case` interface, tidy
`data.frame` returns with stable column schemas, native `ggplot2`
visualization, dependencies that follow current best practice, and a test
suite. The function families have also been broadened, particularly the
`ss_power_*` family, which now covers between-subjects, within-subjects,
mixed, and multi-level designs across the chapters of MDK.

## Why Effect Size Confidence Intervals?

A *p*-value tells you whether an effect is plausibly nonzero in the
direction you tested. It does not tell you how *large* the effect is, and
it does not tell you how *precisely* you have estimated it. Both of those
questions matter scientifically: a study can produce a "significant" but
trivially small effect, or a "nonsignificant" effect whose confidence
interval is so wide that it is consistent with both a small and a large
true value. Reporting an effect size *and* a confidence interval for it
addresses both shortcomings, and the practice is now standard in most
quantitative areas of psychology and education (Kelley & Preacher, 2012).

The technical machinery underneath is less familiar than it should be.
For most standardized effect sizes, the sampling distribution is
*noncentral* (e.g., the noncentral *t* underlying Cohen's $d$, the
noncentral *F* underlying $R^2$, the noncentral $\chi^2$ underlying RMSEA),
which means the correct interval is asymmetric about the point estimate.
The usual symmetric "estimate $\pm$ standard error" shortcut ignores that
skew, so it puts the limits in the wrong places. DMAR computes the
correct intervals from the noncentral distributions directly. This is the
common engine behind every `ci_*()` function in the package.

## Two Philosophies of Sample Size Planning

Researchers who plan studies in DMAR can do so from either of two
complementary perspectives:

  - **Power.** Functions in the `ss_power_*` family answer the question
    *"How many subjects do I need so that, if the effect in the
    population is what I expect, I have at least probability $\pi$ of
    rejecting the null?"* This is the familiar Cohen-style framing.
  - **Accuracy in parameter estimation (AIPE).** Functions in the
    `ss_aipe_*` family answer a different question: *"How many
    subjects do I need so that the confidence interval around my
    effect size estimate is no wider than $w$?"* AIPE is appropriate
    when estimation is the primary goal, for example, when an effect
    is already known to be nonzero and the next study should pin
    down its magnitude rather than re-establish significance.

It is worth being precise about what *accuracy* means in this
framework, because the term is sometimes used loosely. Accuracy
considers precision and bias together: an accurate estimator is one
whose sampling distribution is concentrated *and* centered on the
target parameter. Precision alone, in contrast, can always be
improved by trading bias for precision. A trivial estimator that
returns the same constant value for every sample has zero variance
and is therefore arbitrarily "precise," but it is biased unless the
constant happens to equal the population parameter. The AIPE framing
asks for sample sizes that buy narrow confidence intervals (a
precision target) *without* trading bias to get there. The
confidence interval width target $w$ in `ss_aipe_*` is shorthand for
that joint accuracy goal: a CI that is informative because it is
both narrow and centered on the parameter (Kelley &
Rausch, 2006; Kelley & Maxwell, 2003; Maxwell, Kelley, & Rausch,
2008).

The two perspectives often imply different sample sizes for the
same study. An AIPE plan tends to require more subjects than a
power plan when the expected effect is moderate, because narrow
intervals are more demanding than null rejection. DMAR provides
both side by side so that the choice can be made deliberately, in
light of the scientific question.

## When Do You Reach for DMAR?

The package is organized around five families of tasks. Pick the one that
matches the question you are trying to answer.

### 1. *I Just Ran an ANOVA, t-Test, or Regression. What's the Effect Size and a CI for It?*

Use the `ci_*` family.

```{r}
# Confidence interval for omega squared (proportion of variance accounted
# for) from a one-way ANOVA with F = 11.221 on (4, 50) df, total N = 55.
ci_omega_squared(F_value = 11.221, df_effect = 4, df_error = 50, N = 55)

# More naturally, from a fitted model:
fit <- aov(weight ~ group, data = PlantGrowth)
ci_omega_squared(fit)
```

The output is a tidy `data.frame`: an effect label, the point estimate
(`omega_squared`), the lower and upper limits of the CI, and the inputs
that produced it (`F_value`, `df_effect`, `df_error`, and *N*). That format makes it
easy to feed the result directly into a table or a forest plot
(`plot_ci()` is built to consume it).

CI variants exist for many other effect sizes: `ci_smd()` for the
standardized mean difference (Cohen's $d$); `ci_smd_c()` for the
standardized mean difference using the control-group SD as the divisor;
`ci_pvaf()` for the proportion of variance accounted for by a single
predictor; `ci_snr()` and `ci_srsnr()` for signal-to-noise ratios;
`ci_R()`, `ci_R2()`, and `ci_rc()` for correlations and standardized
regression coefficients; `ci_rmsea()` for the RMSEA model-fit index; and
others.

### 2. *I'm Planning a Study. How Many Subjects Do I Need?*

Use `ss_power_*` for sufficient power, or `ss_aipe_*` for sufficient
precision. The same effect can be approached from either angle.

```{r}
# Power-based: per-group n needed for 90% power to detect a contrast among
# 4 groups with population means (90, 92, 88, 81) and within-group
# variance 144. The contrast compares the average of the first three
# groups against the fourth.
ss_power_contrast(
  c_weights     = c(1/3, 1/3, 1/3, -1),
  mu            = c(90, 92, 88, 81),
  sigma_squared = 144,
  desired_power = 0.90
)
```

```{r}
# AIPE-based: n needed so a 95% CI around a standardized mean difference
# (delta) of 0.50 is no wider than 0.30 units (a tight interval).
ss_aipe_smd(delta = 0.50, conf_level = 0.95, width = 0.30)
```

The `ss_power_*` family in v1.0.0 spans most of the standard designs
covered in MDK: one-way ANOVA (`ss_power_one_way_anova`), factorial ANOVA
(`ss_power_factorial_anova`), repeated measures ANOVA with sphericity
adjustment (`ss_power_rm_anova`), split-plot mixed ANOVA
(`ss_power_split_plot_anova`), Pearson correlation (`ss_power_r`),
two-independent-groups standardized mean difference (`ss_power_smd`),
contrasts standardized and unstandardized (`ss_power_c`, `ss_power_sc`,
`ss_power_c_ancova`, `ss_power_contrast`), regression coefficients
(`ss_power_reg_coef`), and two-level random-intercept models for
cluster-randomized trials (`ss_power_mixed_effects`). For SEM-style
designs, use `ss_power_sem()`. AIPE counterparts exist for the
correspondingly named effects.

### 3. *I Want to Test a Specific Contrast or Set of Pairwise Comparisons.*

Use `contrast_test()` for arbitrary contrasts in a one-way ANOVA, with
optional multiple-comparison adjustment.

```{r}
fit <- aov(weight ~ group, data = PlantGrowth)
# All pairwise comparisons, Tukey-protected family-wise error rate:
contrast_test(fit, contrasts = "pairwise", adjust = "tukey")
```

For the bare critical value of a particular procedure, see the `cv_*`
family: `cv_t()`, `cv_tukey_hsd()`, `cv_scheffe()`, `cv_smm()` (the
studentized maximum modulus), `cv_dunnett()`, and `cv_z()`. These return
the threshold value alone, suitable for inclusion in user code or
hand-checking textbook examples.

### 4. *I Have Within-Subjects (Repeated Measures) Data.*

Start with `anova_within()`, which combines the univariate
within-subjects *F* test with a sphericity diagnostic (Mauchly's *W*) and
the three standard $\varepsilon$-corrected *p*-values
(Greenhouse-Geisser, Huynh-Feldt, lower-bound). The motivation is that an
uncorrected within-subjects *F* test can be substantially anticonservative
when the sphericity assumption fails, and `anova_within()` puts the
diagnostic and the correction in one place.

```{r}
if (requireNamespace("nlme", quietly = TRUE)) {
  res <- anova_within(nlme::Orthodont,
                      id = "Subject", time = "age", outcome = "distance")
  res
}
```

For the underlying components separately, see `mauchly_test()` and
`epsilon_corrections()`. For visualizing individual subject trajectories,
see `plot_trajectories()` and `plot_trajectories_fitted()`.

### 5. *I Want Descriptive Statistics, a Correlation Table, or Distribution-Shape Diagnostics.*

`descriptives()` returns a tidy summary suitable for screening:

```{r}
descriptives(attitude)$descriptives
```

Adding `correlations = TRUE` appends a Pearson correlation matrix. For a
publication-ready correlation table with *p*-values, confidence intervals,
and significance stars (and, optionally, HTML or LaTeX export), use
`correlations_test()`:

```{r}
correlations_test(attitude, stars = TRUE)
```

For interrater agreement, use `cohen_kappa()` (two raters), `fleiss_kappa()`
(multiple raters), or `icc()` (intraclass correlation, all six classical
forms). For distribution-shape statistics in isolation, use `skewness()`
and `kurtosis()`.

## What DMAR Is *Not*

DMAR deliberately overlaps with, but does not duplicate, several
specialized packages.

  - For full-featured factor analysis and SEM, use `lavaan`.
  - For coefficient $\alpha$ / $\omega$ with extensive options, use
    `psych::omega()`.
  - For mixed-effects models, use `lme4` or `nlme`.
  - For robust ANOVA, use `WRS2`.

DMAR layers tidy effect size CIs, sample size planning, and
publication-ready output over those tools, rather than replacing them.

## A Note on Argument Naming

The package uses `snake_case` throughout. Capital letters in identifiers
are reserved for statistically meaningful symbols, for example `R2` for the
squared multiple correlation, `N` for total sample size, `S` for a
covariance matrix, `Lambda` for a factor-loadings matrix, `F_value` for
an *F* statistic. Every other argument is lowercase, with the result
that examples read consistently regardless of discipline:

```{r show-args, eval = FALSE}
ci_R2(R2 = 0.25, N = 100, p = 5, random_predictors = TRUE)
ss_power_reg_coef(rho2_Y_X = 0.78, rho2_Y_X_without_j = 0.74,
                  p = 5, desired_power = 0.85)
ss_aipe_cv_sensitivity(true_cv = 0.25, estimated_cv = 0.25,
                       width = 0.10, conf_level = 0.95, G = 200)
```

The MBESS dot.case style (`conf.level`, `Random.Predictors`,
`Specified.N`) is no longer accepted. Scripts written against MBESS will
need to convert dots to underscores in argument names and lower-case any
prefix words that are not themselves statistical notation. Function
names are unchanged in nearly all cases (`smd()`, `ci_smd()`,
`ci_R2()`, `ss_aipe_reg_coef()`, etc.), and *return* shapes are
unchanged or strictly richer.

If you have scripts written against MBESS or against pre-1.0 DMAR
snapshots and you see errors of the form

```
Error in foo(...) : unused argument (Group.1 = ...)
```

the fix is mechanical: replace dots with underscores in the argument
name (`Group_1`, `n_1`, `Mean_1`, `Specified_N`, etc.) and then
lowercase the non-meaningful capitalized prefix words
(`group_1`, `mean_1`, `specified_N`, ...). See `NEWS.md` for the full
mapping. The exception worth noting on the function side is
`aipe_smd()`, which has been renamed to `ss_aipe_smd()` to align it
with the rest of the `ss_aipe_*` family.

The motivation for the unification is simple: the former mix of
`Random.Predictors` (dot.case for legacy MBESS arguments) and
`conf_level` (snake_case for newer arguments) was confusing in
side-by-side examples, especially in teaching, and a single convention
removes the source of friction.

## Where to Next

- For a worked walkthrough of the visualization functions and
  effect size confidence intervals on a simulated psychometric
  study, see `vignette("effect-size-visualization", package = "DMAR")`.
- For when full information maximum likelihood regression actually
  helps over `lm()`, with `mlmr()` and `mlmr_mv()`, see
  `vignette("mlmr", package = "DMAR")`.
- For the function reference, `?DMAR-package` lists the exported
  functions grouped by family.
- Citation: `citation("DMAR")`.
- Author web site: <https://kenkelley.org>; related publications:
  <https://kenkelley.org/publications/>.

## References

Kelley, K. (2007a). Confidence intervals for standardized effect
sizes: Theory, application, and implementation. *Journal of
Statistical Software, 20*(8), 1--24.
<https://doi.org/10.18637/jss.v020.i08>

Kelley, K. (2007b). Methods for the behavioral, educational, and
social sciences: An R package. *Behavior Research Methods, 39*(4),
979--984. <https://doi.org/10.3758/BF03192993>

Kelley, K., & Maxwell, S. E. (2003). Sample size for multiple
regression: Obtaining regression coefficients that are accurate, not
simply significant. *Psychological Methods, 8*(3), 305--321.

Kelley, K., & Preacher, K. J. (2012). On effect size. *Psychological
Methods, 17*(2), 137--152.

Kelley, K., & Rausch, J. R. (2006). Sample size planning for the
standardized mean difference: Accuracy in parameter estimation via
narrow confidence intervals. *Psychological Methods, 11*(4),
363--385.

Maxwell, S. E., Delaney, H. D., & Kelley, K. (2027). *Designing
experiments and analyzing data: A model comparison perspective*
(4th ed.). Routledge.

Maxwell, S. E., Kelley, K., & Rausch, J. R. (2008). Sample size
planning for statistical power and accuracy in parameter
estimation. *Annual Review of Psychology, 59*, 537--563.
<https://doi.org/10.1146/annurev.psych.59.103006.093735>
