---
title: "Weighted Kappa: Cohen's 1968 Illustration, Worked in Full"
author: "Ken Kelley"
date: "`r format(Sys.Date(), '%B %Y')`"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Weighted Kappa: Cohen's 1968 Illustration, Worked in Full}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
library(DMAR)
```

Cohen's (1968) weighted kappa generalizes his (1960) kappa to
situations in which some disagreements are graver than others: each
cell of the k x k agreement matrix receives a weight, and the
coefficient is the chance-corrected proportion of weighted agreement.
This vignette works through Cohen's own illustration from start to
finish using `cohen_kappa()` and the `diagnosis_agreement` data set,
which reproduces his Table 1 cell for cell. Along the way it covers
the two weight scalings, the standard error question, linear and
quadratic weighting for ordinal categories, Cohen's connection between
weighted kappa and the product-moment correlation, and, at the end, an
orientation question about the paper's validity example that the
reader can adjudicate with all computations in view.

## The Data: Cohen's Table 1

Two judges independently assign *N* = 200 cases to three
diagnostic categories. The `diagnosis_agreement` data set stores one
row per cell of the 3 x 3 matrix, in the paper's own layout (Judge B
indexes the rows, Judge A the columns), with the three quantities
Cohen prints in each cell: the ratio-scaled disagreement weight, the
chance-expected proportion (his parenthetical values), and the
observed proportion.

```{r}
data(diagnosis_agreement)
diagnosis_agreement

tab <- xtabs(frequency ~ judge_b + judge_a, data = diagnosis_agreement)
addmargins(tab)
```

The margins are .50/.30/.20 for Judge B (rows) and .60/.30/.10 for
Judge A (columns), and each cell's expected proportion is the product
of its row and column marginal proportions:

```{r}
with(diagnosis_agreement,
     all.equal(expected_proportion,
               as.numeric(tapply(observed_proportion, judge_b, sum)[judge_b] *
                          tapply(observed_proportion, judge_a, sum)[judge_a])))
```

## Unweighted Kappa

With every disagreement treated alike, kappa is the proportion of
agreement corrected for chance (Cohen, 1960):

```{r}
cohen_kappa(table = tab)
```

The estimate is .492: after chance agreement is removed from both the
numerator and the base, about half the judgments agree. The confidence
interval uses the Fleiss, Cohen, and Everitt (1969) asymptotic
standard error.

## Disagreement Scaling and Weighted Kappa

Cohen scales *disagreement*: the diagonal gets 0, a personality
disorder-neurosis confusion gets 1, personality disorder-psychosis
gets 3, and neurosis-psychosis, the confusion the illustration treats
as gravest, gets 6. The weights are part of the definition of
agreement, so they must be fixed before the data are collected, and
weighted kappa is invariant to multiplying them by any positive
constant.

```{r}
v <- unclass(xtabs(disagreement_weight ~ judge_b + judge_a,
                   data = diagnosis_agreement))
v

res <- cohen_kappa(table = tab, weights = v,
                   weight_scaling = "disagreement")
res
```

Weighted kappa is .348, *smaller* than the unweighted .492.
Cohen chose these values to make a point that is easy to miss: because
the same weights enter both the observed and the chance-expected
weighted disagreement, weighted kappa is fully chance corrected, and
"partial credit" does not push it upward. These judges disagree far
less than chance expectation in the mildly weighted cells and at about
the chance level in the heavily weighted ones; they disagree least
where it matters least. Interchanging the 6 and 1 weights reverses the
conclusion:

```{r}
v_swap <- v
v_swap[v == 6] <- 1
v_swap[v == 1] <- 6
cohen_kappa(table = tab, weights = v_swap,
            weight_scaling = "disagreement")
```

The per-cell quantities behind any result travel with it:

```{r}
attr(res, "cells")
```

## Cohen's Standard Errors and the Modern Interval

Cohen's Formulas 10 and 13 give large-sample standard errors for
weighted kappa; the arithmetic below reproduces his printed values
from the data set's columns.

```{r}
q_o <- with(diagnosis_agreement,
            sum(disagreement_weight * observed_proportion))
q_c <- with(diagnosis_agreement,
            sum(disagreement_weight * expected_proportion))
c(q_o = q_o, q_c = q_c, kappa_w = 1 - q_o / q_c)

v2_o <- with(diagnosis_agreement,
             sum(disagreement_weight^2 * observed_proportion))
v2_c <- with(diagnosis_agreement,
             sum(disagreement_weight^2 * expected_proportion))
c(se_formula_10 = sqrt((v2_o - q_o^2) / (200 * q_c^2)),
  se_formula_13 = sqrt((v2_c - q_c^2) / (200 * q_c^2)))
```

Those are his .0901 and .0916, his 95% interval is .348 plus or minus
1.96 times .0901, and his significance test is z = .348/.0916 = 3.80.
A year later, Fleiss, Cohen, and Everitt (1969) derived the standard
error now in standard use, which `cohen_kappa()` reports; on these
data it is noticeably smaller (.0755), so the modern interval is
tighter than the one printed in the 1968 paper.

## Linear and Quadratic Weights for Ordinal Categories

When categories are ordered, two standard weight patterns need no
custom matrix: `weights = "linear"` and `weights = "quadratic"`. Both
are stated on the agreement scale internally, and the quadratic
pattern is the agreement-scale counterpart of disagreement weights
proportional to the squared category distance, $(i - j)^2$.

```{r}
df <- as.data.frame(as.table(tab))
score_b <- rep(as.integer(df$judge_b), df$Freq)
score_a <- rep(as.integer(df$judge_a), df$Freq)

v_quad <- outer(1:3, 1:3, function(i, j) (i - j)^2)
cohen_kappa(table = tab, weights = v_quad,
            weight_scaling = "disagreement")$kappa
cohen_kappa(score_b, score_a, weights = "quadratic")$kappa
```

Cohen (1968) proves that with quadratic-pattern weights and equal
marginal distributions, weighted kappa *equals* the
product-moment correlation between the category scores, which offers
a useful way to think about what r is: a chance-corrected proportion
of agreement with disagreements weighted by squared distance. With
the unequal marginals of Table 1 the two are close but not equal,
with weighted kappa the smaller, as Cohen notes:

```{r}
c(kappa_w_quadratic = cohen_kappa(score_b, score_a,
                                  weights = "quadratic")$kappa,
  pearson_r = cor(score_b, score_a))
```

On a table with equal marginals the identity is exact:

```{r}
eq <- matrix(c(40, 10, 10,
               10, 60, 10,
               10, 10, 40), nrow = 3, byrow = TRUE)
dfe <- as.data.frame(as.table(as.table(eq)))
t1 <- rep(as.integer(dfe$Var1), dfe$Freq)
t2 <- rep(as.integer(dfe$Var2), dfe$Freq)
c(kappa_w_quadratic = cohen_kappa(table = eq, weights = v_quad,
                                  weight_scaling = "disagreement")$kappa,
  pearson_r = cor(t1, t2))
```

The quadratic pattern is also the choice under which weighted kappa is
equivalent to the intraclass correlation (Fleiss & Cohen, 1973), which
is why it is the most common weighting for ordinal scales.

## The Validity Example and an Orientation Question

Cohen closes the paper by reinterpreting Table 1 as a validity
problem: Judge A becomes the consensus diagnosis of a panel (the
criterion) and Judge B a computer diagnosis (the predictor). Because a
predictor's two directions of confusion can carry different costs,
the weights may be asymmetric. His prose says a computer diagnosis of
neurosis when the panel consensus is psychosis (weight 6) is graver
than a computer psychosis when the panel says neurosis (weight 2),
and the weight display printed in the paper, which has Panel labeling
its columns and Computer its rows, matches that prose:

```{r}
v_printed <- matrix(c(0, 1, 4,
                      1, 0, 6,
                      2, 2, 0), nrow = 3, byrow = TRUE,
                    dimnames = list(computer = c("D", "N", "P"),
                                    panel    = c("D", "N", "P")))
v_printed
```

Here is the question. Table 1 has Judge B (the computer, in this
reinterpretation) on its rows and Judge A (the panel) on its columns,
the same orientation as the printed weight display, so applying the
printed weights cell by cell gives one answer; applying their
transpose gives another. Cohen reports the weighted disagreement sums
as .86 (observed) and 1.33 (chance), yielding weighted kappa .353,
with Formula 10 and 13 standard errors .0887 and .0915. Both
computations follow, from the same data:

```{r}
# As printed, read with its own row and column labels:
res_printed <- cohen_kappa(table = tab, weights = v_printed,
                           weight_scaling = "disagreement")
cells <- attr(res_printed, "cells")
c(sum_v_p_obs = sum(cells$disagreement_weight * cells$observed_proportion),
  sum_v_p_exp = sum(cells$disagreement_weight * cells$expected_proportion),
  kappa_w     = res_printed$kappa)

# The transpose of the printed display:
res_t <- cohen_kappa(table = tab, weights = t(v_printed),
                     weight_scaling = "disagreement")
cells_t <- attr(res_t, "cells")
c(sum_v_p_obs = sum(cells_t$disagreement_weight * cells_t$observed_proportion),
  sum_v_p_exp = sum(cells_t$disagreement_weight * cells_t$expected_proportion),
  kappa_w     = res_t$kappa)
```

The transpose reproduces Cohen's published quantities (.86, 1.33,
.353, and his Formula 10 and 13 values .0887 and .0915 follow from
the same sums); the printed orientation gives .62, 1.07, and .421
instead. So the published numerical results correspond to the weights
applied in the transpose of the orientation the printed display's
labels state. What to make of that is a judgment the numbers alone do
not settle: the possibilities include a transposed weight display in
typesetting, a transposed data table relative to the notation used in
computation, or a slip in the computation itself, and nothing in the
paper distinguishes among them. Cohen's internal arithmetic is
consistent throughout, so the substance of the example, that
asymmetric weights make weighted kappa a validity measure, is
untouched either way. The `cohen_kappa()` help page follows the
published values (weighted kappa .353), which is the analysis the
paper reports.

## References

Cohen, J. (1960). A coefficient of agreement for nominal scales.
*Educational and Psychological Measurement, 20*(1), 37--46.

Cohen, J. (1968). Weighted kappa: Nominal scale agreement provision
for scaled disagreement or partial credit. *Psychological
Bulletin, 70*(4), 213--220.

Fleiss, J. L., & Cohen, J. (1973). The equivalence of weighted kappa
and the intraclass correlation coefficient as measures of
reliability. *Educational and Psychological Measurement, 33*(3),
613--619.

Fleiss, J. L., Cohen, J., & Everitt, B. S. (1969). Large sample
standard errors of kappa and weighted kappa. *Psychological
Bulletin, 72*(5), 323--327.
